Exploring the importance of a complete statistical description of dependence in ecology and evolution

### 2018-02-14 All_analysis.Rmd
mostly changes to implement the bivariate models selection on Terry's C/N data, and to randomize copula structure and see what it does to the distribution of some ecosystem function

@@

None of the copulas we tried here was a good fit, but some are much 
better than others, and the normal copula is not close to the best, 
so that shows there is non-trivial (i.e., non-normal) copula structure.

It looks like the data might be asymmetric about the y=x line, which
might require a special copula to model. What is the significance,
biologically?

Strong dependence in the extreme right tails is evident. We could
quantify this, and (probably) show it is significantly more than what one would expect from a null hypothesis of a normal copula with the same kendall as the data. What is the significance, biologically?

>>>>

<!--A function needed for caching-->
```{r echo=F}
mtime <- function(files)
{
  lapply(Sys.glob(files),function(x) X =   
          file.info(x)$mtime)
}
```

@@

None of the copulas we tried here was a good fit, but some are much 
better than others, and the normal copula is not close to the best. 
So that shows there is non-trivial (i.e., non-normal) copula structure, which is an interesting conclusion. It is more interesting than it would have been to specifically identify a copula that fits, so I am not too concered about our failure to do that.

It looks like the data might be asymmetric about the y=x line, which
might require a special copula to model. What is the significance of this,
biologically?

Strong dependence in the extreme right tails is evident. We could
quantify this, and (probably) show it is significantly more than what one would expect from a null hypothesis of a normal copula with the same kendall as the data. What is the significance, biologically? We could (probably) also show there is much more asymmetry of tail dependence (upper v. lower) than would be expected from a normal-copula null hypothesis.

Probably the most promising next step is to get a function $f(N,C)$ from Amy or Terry (or someone else) which is the expected value of some ecosystem function (or other quantity of interest), supported by a publication. Then calculate its value on all the sites and look at the distribution and calculate, say, its skewness or some other property. Then get surrogate, randomized data with normal copula structure (or some other copula structure - try Clayton because it has the reverse tail dependence). Then calculate the distribution of $f(N,C)$ for each surrogate dataset and its skewness (or other statistic) and compare. I already have code for this (`copsurrog2d.R`). So let's try this. Here is a function:

```{r ecofunc_Loecke, cache=T, cache.extra=list(d,mtime("copsurrog2d.R"))}
#Expected value of some ecosystem function for a 
#soil based on soil N and C
#
#Args
#N    Vector of N values for a bunch of soils
#C    Vector (same length as N) of C values for the same soils
#
#Output
#Expected value of some ecosystem function for those soils (vector of the same length as N and C).
#
ecofunc<-function(N,C)
{
  #right now this is just a placeholder, a real formula and a specific ecosystem function of interest to be specified by Amy and Terry
  return(N+C)
}

source("copsurrog2d.R")
source("SkewnessAnd3CentMom.R")
inds<-which(is.finite(d[,1]) & is.finite(d[,2]))
d<-d[inds,]

#compare to a normal-copula null
cop<-normalCopula(.5,2)
numresamp<-1000
dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=numresamp)
efval<-myskns(ecofunc(d[,2],d[,1]))

efvalsurN<-NA*numeric(numresamp)
for (counter in 1:numresamp)
{
  efvalsurN[counter]<-myskns(ecofunc(dsur[,2,counter],dsur[,1,counter]))
}
sum(efvalsurN<efval)/numresamp
hist(efvalsurN,50)
points(efval,0,col='red')

#compare to a clayton-copula null
cop<-claytonCopula(.5,2)
numresamp<-1000
dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=numresamp)
efvalsurC<-NA*numeric(numresamp)
for (counter in 1:numresamp)
{
  efvalsurC[counter]<-myskns(ecofunc(dsur[,2,counter],dsur[,1,counter]))
}
sum(efvalsurC<efval)/numresamp
hist(efvalsurC,50)
points(efval,0,col='red')

#compare to a joe-copula null
cop<-joeCopula(3,2)
numresamp<-1000
dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=numresamp)
efvalsurJ<-NA*numeric(numresamp)
for (counter in 1:numresamp)
{
  efvalsurJ[counter]<-myskns(ecofunc(dsur[,2,counter],dsur[,1,counter]))
}
sum(efvalsurJ<efval)/numresamp
hist(efvalsurJ,50)
points(efval,0,col='red')
```
